---
title: "R for Data Science With Sports Applications"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```


## Recap

- From Lab 1: it is important to keep track of where files are located in your computer.

- Your hard drive is organized in _folders_ or _directories_.

- In Mac os, `~/Desktop` is your desktop. 

- If you have a directory called lab 1 in your desktop, `~/Desktop/lab1` is the location of that folder.

- Run `setwd("~/Desktop/lab1")` to make that the working directory.

- `list.files()` prints the contents of the working directory in your console.


---
## Recap 2

- R code is organized in functions.

- Functions take arguments and return values.

- Data is stored in objects.

- Assignment (`<-`) makes variables point to objects.

---
## Recap 3

```{r}
x <- c(1, 2, 3)
mean(x)
```

```{r}
str(mtcars)
```

---
## Data frame

- The most important kind of object in R.

```{r}
# NBA championship 2017/2018 82 regular season games
df <- readRDS('./nba.rds')
```

---
## Inspect df
```{r}
df
```

---
```{r}
library(dplyr)
glimpse(df)
```

---
## Visualize

We want to understand the relationship between assists and turnovers.

```{r, echo=FALSE}
library(ggplot2)


# TODO: create a plot of assists vs turnovers.
ggplot(data=df, aes(x=AST, 
                    y=TOV, 
                    color=Playoff, 
                    shape=Playoff)) + 
  geom_point() + 
  labs(title="Turnovers vs. Assistes",
       subtitle="2017/18 Regular Season",
       x="Assists",
       y="Turnovers")
  
```

---
## Grammar of graphics

- Data

- Aesthetic mappings

- Geoms


---
### Identify the components of the grammar of graphics


```{r}
library(trackeR)
runZones <- zones(runs[1:4], what = "speed", breaks = c(0, 2:6, 12.5))
plot(runZones)
```

---
### Lab 2

- Create a plot to understand the relationship between two pointers made and three pointers made.
- Think about the data and the aesthetic mappings that you want.


